The Two Dualities of Computation: Negative and Fractional Types

### 2012-02-28 rational.tex
notes on multiplicative duals

%%%%%%%%%%%%%%%%%%%%%
\subsection{Multiplicative Duality in {{langRev}} }

@@

@@

\paragraph*{Unification.}

test test test
\begin{multicols}{2}
\begin{center}
\scalebox{1.5}{
  \includegraphics{diagrams/eta_times1.pdf}
}
\end{center}

@@

\paragraph*{Entanglement.}

\paragraph*{Coherence.}

\paragraph*{Annihilation.}

\paragraph*{Zero Information.}

@@

%%%%%%%%%%%%%%%%%%%
\subsection{SAT solver}

>>>>

ADD MATH.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Multiplicative Duality in {{langRev}} }

@@

As with additive duals, we represent {{eta*}} and {{eps*}] using
`U'-shaped connectors. In the general case the wires for {{1}} will be
dropped as it can freely introduced or eliminated using {{uniti}} and
{{unite}}.

@@

The usual interpretation of {{b1*b2}} we have both a value of type
{{b1}} and a value of type {{b2}}. This interpretation is maintained
in the precense of fractionals. Hence an {{eta* : 1 <-> b^^^ * b}} is
to be viewed as a fission point for a value of type {{b}} and its
multiplicative inverse {{b^^^}}. This naturally raises the question --
which value of type {{b}} is produced? The isomorphism {{eta*}} does
not produce any specific value of type {{b}}, instead it produces a
place holder for a value such that the value of type {{b^^^}} is
always the dual of the chosen one.

\begin{multicols}{2}
\begin{center}
\scalebox{1.5}{
  \includegraphics{diagrams/eta_times1.pdf}
}
\end{center}

@@

\paragraph*{Unification.}
A computational interpretation of this is that {{langRevEE}} acts like
a logic programming system and {{eta*}} acts as the site of creation
of a new typed logic variable of type {{b}} and its corresponding dual
of type {{b^^^}}. These two values are free to move independantly
through a circuit, but any operations on value affect what operations
are possible on the other since they are both related by an underlying
logic variable. Dually, {{eps*}} acts as a unification site for values
flowing along both branches. If the values are not exactly duals, then
the system fails to unify and hence the program is stuck. If the
values are exactly duals, then unification succeeds and {{eta*}}
returns {{1}}.

\paragraph*{Entanglement and Anti-Particles.}
A dual view of the operation of {{eta*}} and {{eps*}} is inspired by
quantum mechanics. The {{eta*}} is a site of fission that creates an
information particle and its anti-particle. Both flow in the same
direction in time. The produce particles are entangles in that actions
on one (such as transformation by application of an isomorphism)
affects the other. Further the fission site does not create a particle
in any specific state, but in a superposition of all possible states
as dictated by its type. In each possible world that the particle
exists, it takes on a specific value inhibiting its type and
correspondingly its entangled anti-particle takes the dual of the
specific value.

This also leads to the exiciting point of view that entangled
particle/anti-particle pairs act as functions in a physical
first-class sense. We can ask the fuction to be applied to a value by
unifying the anti-particle with a particle in a well-known state and
examining is pair. We shall use this point of view in the development
our SAT solver (see Sec. \ref{sec:sat-solver}).

\paragraph*{Coherence.} 
As before, the requisite coherence condition on Compact Closed
Categories is applicable and is expressed below in terms of
{{eta*}}/{{eps*}} and the multiplicative monoid.

\paragraph*{Annihilation.}
As with {{(+, 0)}} we can construct a {{trace*}} operation. Unlike the
additive {{trace+}}, {{trace*}} gives us the ability to find the
fixpoint of the contraint (expressed with combinators) that is
traced. If the constraint has not satisfying values, there is not
fixpoint possible -- we use the term `annihilation' to describe the
corresponding undefined state of the program. Annihilation and a
non-termination both describe undefined program execution -- while
non-termination characterizes an iteration that fails to terminate,
annihilation characterizes a constraint that has no fixpoint.

\paragraph*{Zero Information.}
As with {{eta+}}, we argue that {{eta*}} constructs no new
information, instead it gives us the ability to explore as many worlds
as permitted by the type {{b}}. So situate ourselves in any specific
world, we have to supply additional information in teh form of
unifying the particle or anti-particle with a particle representing
known information. Hence only by supplying information to a an
entangled pair can we read information from it -- and always exactly
as much information as we supplied.

ADD MATH.

@@

%%%%%%%%%%%%%%%%%%%
\subsection{SAT solver}
\label{sec:sat-solver}
